### 2021-02-12 manuscript.tex
converted bibtex style to zotero and updated cite keys

\subsection{Server Setup}
At it's core, the OS transition from RHEL 6 to 7 was the original project requirement that spurred this work. Existing hosting for the CCGD was owned by university OIT who established an end of life date for the current server OS\@. As a result, the bulk of core architecture implementation followed a narrow prescription for what was termed a ``Fully Managed'' RHEL7 install as detailed by the OIT-LPT Public-Docs~\cite{OIT}. Contrary to the implications of this installation name, a Fully Managed administration carried a mixture of privileges and management roles that were shared between OIT and our group. This was accomplished by leveraging a chef admin paradigm that was allowed limited sudo privileges within the hosted ecosystem. Importantly, items that extend from the core of server administration such as virtual machine creation, backups, and core security were handled by OIT\@. This left our group with the more focused task of administering exclusively our application.

@@

The structure of a general shiny app follows a highly reduced and consistent paradigm, and ours is no different. The broad format of this work flow is \texttt{ui + server $\rightarrow$ deployment}. The server specifically has a discrete composition made up of reactive constructs of inputs, expressions, and outputs that shape the app's reactive graph, which forms, more accurately, the discrete operation of the application. The assembly and movement of this graph is both emblematic and practical in displaying the modularity and improved resource utilization of the reactive elements. Without belaboring the description of reactive programming, this model employs a declarative programming style that allows for lazy code execution~\cite{Wickham}. What this amounts to in application is for the developer to define loose controls at a higher level while allowing the code to passively fill these requirements. This translates to greater flexibility to both the developer and the machine as resource use can be reduced through modularity in code execution.

@@

\subsection{Best Practices}
Some features of the software best practices from the previous version of the CCGD were done quite well such as the level of detail in the documentation. At the same time, many other important areas were missing in the development and maintenance of the application. Most prominent among these were the lack of any degree of version control and relatively inconsistent design principles. As a result, our group employed or improved upon several methods in an effort to elevate the quality, usability, and transportability of our codebase and application~\cite{Fincham2011}.

@@

In approaching the process of development, we applied a philosophy of tidy code creation to both syntax and design. A supporting method in creating a more elegant code syntax was the use of R universe tools, Tidyverse and Rmarkdown. Both offer a coding grammar that is centered around accessibility and simplification of high-level language organization~\cite{Wickham2019}. Furthering this simplification, we sought opportunities to deduplicate and extract subroutines in the table integration process in an effort to streamline resource constraints. Finally, another notable area for improvement we sought was in disk space utilization. The existing integration process relied on the continuous presence of what were several intermediary files that occupied a sizable disk capacity. To combat this we specifically reworked several data transformation steps to allow for the post-process elimination of these temporary data objects.

@@

\subsection{Architecture Upgrades}
Following the goals set out in our project requirements, we chose Rmarkdown as a replacement framework to develop the application's web front-end. Previously, the web interface for the application had been developed using Adobe Dreamweaver~\cite{Abbott2015} resting on a traditional LAMP stack. For the time, a LAMP stack and Dreamweaver served the purpose of simplifying the process of interface construction on a stable and consistent web platform, yet we hoped to extend the theme of simplified content creation into a modern framework. Further, utilizing a language such as Rmarkdown almost fully carries the development process into a state of pure content creation. Through the Rmarkdown toolset, the work of web design is now largely rendered automatically for the developer (see Figure~\ref{fig:renderFlow}). Employing the R web scheme established a cornerstone for the project to leverage both enhanced web development as well as making available other key tools in architecture design choices.

>>>>

\subsection{Server Setup}
At it's core, the OS transition from RHEL 6 to 7 was the original project requirement that spurred this work. Existing hosting for the CCGD was owned by university OIT who established an end of life date for the current server OS\@. As a result, the bulk of core architecture implementation followed a narrow prescription for what was termed a ``Fully Managed'' RHEL7 install as detailed by the OIT-LPT Public-Docs~\cite{oitOITLPTPublicDocsThis}. Contrary to the implications of this installation name, a Fully Managed administration carried a mixture of privileges and management roles that were shared between OIT and our group. This was accomplished by leveraging a chef admin paradigm that was allowed limited sudo privileges within the hosted ecosystem. Importantly, items that extend from the core of server administration such as virtual machine creation, backups, and core security were handled by OIT\@. This left our group with the more focused task of administering exclusively our application.

@@

The structure of a general shiny app follows a highly reduced and consistent paradigm, and ours is no different. The broad format of this work flow is \texttt{ui + server $\rightarrow$ deployment}. The server specifically has a discrete composition made up of reactive constructs of inputs, expressions, and outputs that shape the app's reactive graph, which forms, more accurately, the discrete operation of the application. The assembly and movement of this graph is both emblematic and practical in displaying the modularity and improved resource utilization of the reactive elements. Without belaboring the description of reactive programming, this model employs a declarative programming style that allows for lazy code execution~\cite{wickhamMasteringShinyBook}. What this amounts to in application is for the developer to define loose controls at a higher level while allowing the code to passively fill these requirements. This translates to greater flexibility to both the developer and the machine as resource use can be reduced through modularity in code execution.

@@

\subsection{Best Practices}
Some features of the software best practices from the previous version of the CCGD were done quite well such as the level of detail in the documentation. At the same time, many other important areas were missing in the development and maintenance of the application. Most prominent among these were the lack of any degree of version control and relatively inconsistent design principles. As a result, our group employed or improved upon several methods in an effort to elevate the quality, usability, and transportability of our codebase and application~\cite{finchamSoftwareDevelopmentPractices2011}.

@@

In approaching the process of development, we applied a philosophy of tidy code creation to both syntax and design. A supporting method in creating a more elegant code syntax was the use of R universe tools, Tidyverse and Rmarkdown. Both offer a coding grammar that is centered around accessibility and simplification of high-level language organization~\cite{wickhamWelcomeTidyverse2019}. Furthering this simplification, we sought opportunities to deduplicate and extract subroutines in the table integration process in an effort to streamline resource constraints. Finally, another notable area for improvement we sought was in disk space utilization. The existing integration process relied on the continuous presence of what were several intermediary files that occupied a sizable disk capacity. To combat this we specifically reworked several data transformation steps to allow for the post-process elimination of these temporary data objects.

@@

\subsection{Architecture Upgrades}
Following the goals set out in our project requirements, we chose Rmarkdown as a replacement framework to develop the application's web front-end. Previously, the web interface for the application had been developed using Adobe Dreamweaver~\cite{abbottCandidateCancerGene2015a} resting on a traditional LAMP stack. For the time, a LAMP stack and Dreamweaver served the purpose of simplifying the process of interface construction on a stable and consistent web platform, yet we hoped to extend the theme of simplified content creation into a modern framework. Further, utilizing a language such as Rmarkdown almost fully carries the development process into a state of pure content creation. Through the Rmarkdown toolset, the work of web design is now largely rendered automatically for the developer (see Figure~\ref{fig:renderFlow}). Employing the R web scheme established a cornerstone for the project to leverage both enhanced web development as well as making available other key tools in architecture design choices.
